---
title: "Ch's 6 - Data Management"
author: "**Yigit Aydede - RBootcamp - 2022**"
output:
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(nycflights13)
library(Lahman)
library(RBootcamp)
library(ggplot2)
tutorial_options(
  exercise.timelimit = 600,
  # A simple checker function that just returns the message in the check chunk
  exercise.checker = function(check_code, ...) {
    list(
      message = eval(parse(text = check_code)),
      correct = logical(0),
      type = "info",
      location = "append"
    )
  }
)
knitr::opts_chunk$set(error = TRUE)
```

## `dplyr`

`dplyr` is a grammar of data manipulation, providing a consistent set of verbs that help you solve the most common data manipulation challenges:

`mutate()` adds new variables that are functions of existing variables
`select()` picks variables based on their names.
`filter()` picks cases based on their values.
`summarise()` reduces multiple values down to a single summary.
`arrange()` changes the ordering of the rows.

See <https://dplyr.tidyverse.org> for more details.

## `filter()`
Use `ames` and see its details at `?ames`.  Explore the data using `glimps()`

```{r m1, exercise=TRUE}

```

Create a new dataframe named `amdf1` that contains 1 or 2-story houses that are built before year 2000 (not including 2000), sold on or after year 2009.

```{r m2, exercise=TRUE}
amdf1 <- filter(ames, Year_Built < 2000,
                Year_Sold >= 2009,
                House_Style == "Two_Story" | House_Style == "One_Story")
```

Find all houses that are sold in 2007, remodeled in 2006, and with with 2 or 3 bedrooms.  Create `amdf2`

```{r m3, exercise=TRUE}
amdf2 <- filter(ames, Year_Remod_Add < 2006,
                Year_Sold == 2007,
                TotRms_AbvGrd == 3 | TotRms_AbvGrd == 2)
```


## `arrange()`

Find the top 10 2-story houses built in 2009 with the highest sale price.  Store these 10 observations in `amdf3`.  Hint use `%>%`

```{r m3a, exercise=TRUE}
temp <- ames %>%
  filter(Year_Built == 2006, House_Style == "Two_Story") %>%
  arrange(desc(Sale_Price))

amdf3 <- temp[1:10,]                
```

Represent the distribution of kitchen quality with a bar chart (use `ggplot2`) among all houses built in 2008 by the ascending order of sold year. Combine all operations by using the pipe operator.

```{r m4, exercise=TRUE}
ames %>%
  filter(Year_Built == 2008) %>%
  arrange(Year_Sold) %>%
  ggplot(mapping = aes(x = Gr_Liv_Area, y = Sale_Price)) + 
  geom_point()
```

Represent the distribution of kitchen quality with a bar chart (use `ggplot2`) among all houses built in 2008 by the ascending order of sold year. Combine all operations by using the pipe operator.

```{r m5, exercise=TRUE}
ames %>%
  filter(Year_Built == 2008) %>%
  arrange(Year_Sold) %>%
  ggplot(mapping = aes(x = Kitchen_Qual)) + 
  geom_bar()
```

Use `subset()` find all houses that are sold in 2007, remodeled in 2006, and with with 2 or 3 bedrooms.

```{r m6, exercise=TRUE}
subset(ames, Year_Sold == 2007  | Year_Remod_Add < 2006 &
       TotRms_AbvGrd == 3 | TotRms_AbvGrd == 2)
```

## `select()`

Select all the variables with names ending with "Qual".

```{r m7, exercise=TRUE}
head(select(ames, ends_with("Qual"))) 
```

Select all the variables with names that contain with "mo".

```{r m8, exercise=TRUE}
head(select(ames, contains("mo"))) 
```

Do not use `dplyr` and try to answer the previous question by using `greb()`

```{r m9, exercise=TRUE}
ind <- grep("mo", colnames(ames))
head(ames[,ind])
```

## `mutate()`, `group_by()` & `summarize()`

Summarize the following values for 2-story & 1-story houses built between 2007 and 2009 with the highest sale price.  
 
1. Number of houses
2. Average number of rooms
3. The mean value of the top 10 sale price 

```{r m10, exercise=TRUE}
ames %>%
  filter(Year_Built >= 2007 & Year_Built <= 2009, 
         House_Style == "Two_Story" | House_Style == "One_Story") %>%
  arrange(desc(Sale_Price)) %>%
  summarize(n_houses = n(),
            ave_n_rooms = mean(TotRms_AbvGrd),
            ave_price = mean(Sale_Price[1:10])) 
```

Do the same summary by kitchen quality

```{r m11, exercise=TRUE}
ames %>%
  filter(Year_Built >= 2007 & Year_Built <= 2009, 
         House_Style == "Two_Story" | House_Style == "One_Story") %>%
  arrange(desc(Sale_Price)) %>%
  group_by(Kitchen_Qual) %>%
  summarize(n_houses = n(),
            ave_n_rooms = mean(TotRms_AbvGrd),
            ave_price = mean(Sale_Price[1:10])) 
```

Add a variable, `ave_lot`, that reflects average lot area by house style in a new data.frame `dff`. 

```{r m12, exercise=TRUE}
dff <- ames %>%
  group_by(House_Style) %>%
  mutate(ave_lot = mean(Lot_Area))
```




