---
title: "Ch's 3 - Other R Objects"
author: "**Yigit Aydede - RBootcamp - 2022**"
output:
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(nycflights13)
library(Lahman)
tutorial_options(
  exercise.timelimit = 60,
  # A simple checker function that just returns the message in the check chunk
  exercise.checker = function(check_code, ...) {
    list(
      message = eval(parse(text = check_code)),
      correct = logical(0),
      type = "info",
      location = "append"
    )
  }
)
knitr::opts_chunk$set(error = TRUE)
```


## Creating a matrix

R stores matrices (and arrays) in a similar way as vectors, but with the attribute called dimension. A matrix is an array that has two dimensions. Data in a matrix are organized into rows and columns. 

1. Create 3 vectors each has 5 numbers randomly selected from 1, 2,. . . ,10.  Name each vector properly
2. Create row matrix from these 3 vectors, called it `A`
3. Create column matrix from these 3 vectors, called it `B`
4. Calculate means of each row and column of `A`.  Assign them a vector, respectively.
5. Assign `V1, V2, ...` to `B` as column names, check it    

```{r m1, exercise=TRUE}

```

## Subsetting matrix

1. Create a 5x5 matrix, `X` sampled from 1:10 with replacement (`set.seed(50)`)
2. Create another matrix from `X` by selecting the 2nd and the 5th rows.  Check it
3. What are the numbers in row 2 and columns 3 and 5

```{r m3, exercise=TRUE}

```

1. What are the numbers in a row when the number in column 3 is 1?
2. Change 10's to 0's
3. Change 0's in the 5th column to 11

```{r m4-setup}
set.seed(50)
X <- matrix(sample(1:10, 25, replace = T), 5,5)
```

```{r m4, exercise=TRUE}

```

1. Find the index (row, column) where 1 is located in `X`.  Hint use `which()`
2. Multiply it with 0.5

```{r m5-setup}
set.seed(50)
X <- matrix(sample(1:10, 25, replace = T), 5,5)
```

```{r m5, exercise=TRUE}

```

Create a 6 × 10 matrix of random integers chosen from 1, 2,. . . ,10.  Call it `aMat` (with `set.seed(75)`)
  
1. Find the number of entries in each row which are greater than 4. You can use `apply()` function. 
2. Which rows contain exactly two occurrences of the number seven? 


```{r m6, exercise=TRUE}

```






*Here's an exercise with some prepopulated code as well as `exercise.lines = 5` to provide a bit more initial room to work.*

Now write a function that adds any two numbers and then call it:

```{r add-function, exercise=TRUE, exercise.lines = 5}
add <- function() {
  
}
```

## Topic 2

### Exercise with Hint

*Here's an exercise where the chunk is pre-evaluated via the `exercise.eval` option (so the user can see the default output we'd like them to customize). We also add a "hint" to the correct solution via the chunk immediate below labeled `print-limit-hint`.*

Modify the following code to limit the number of rows printed to 5:

```{r print-limit, exercise=TRUE, exercise.eval=TRUE}
mtcars
```

```{r print-limit-hint}
head(mtcars)
```

### Quiz

*You can include any number of single or multiple choice questions as a quiz. Use the `question` function to define a question and the `quiz` function for grouping multiple questions together.*

Some questions to verify that you understand the purposes of various base and recommended R packages:

```{r quiz}
quiz(
  question("Which package contains functions for installing other R packages?",
    answer("base"),
    answer("tools"),
    answer("utils", correct = TRUE),
    answer("codetools")
  ),
  question("Which of the R packages listed below are used to create plots?",
    answer("lattice", correct = TRUE),
    answer("tools"),
    answer("stats"),
    answer("grid", correct = TRUE)
  )
)
```

