---
title: "Chapter 3 - Other R Objects: Matrix & Dataframe"
author: "**Yigit Aydede - RBootcamp - 2022**"
output:
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(nycflights13)
library(Lahman)
library(RBootcamp)
tutorial_options(
  exercise.timelimit = 600,
  # A simple checker function that just returns the message in the check chunk
  exercise.checker = function(check_code, ...) {
    list(
      message = eval(parse(text = check_code)),
      correct = logical(0),
      type = "info",
      location = "append"
    )
  }
)
knitr::opts_chunk$set(error = TRUE)
```


## Creating a matrix

R stores matrices (and arrays) in a similar way as vectors, but with the attribute called dimension. A matrix is an array that has two dimensions. Data in a matrix are organized into rows and columns. 

1. Create 3 vectors each has 5 numbers randomly selected from 1, 2,... ,10.  Name each vector properly.
2. Create row matrix from these 3 vectors, called it `A`
3. Create column matrix from these 3 vectors, called it `B`
4. Calculate means of each row and column of `A`.  Assign them separate vectors.
5. Assign `V1, V2, ...` to `B` as column names, check it    

```{r m1, exercise=TRUE}
#1
x <- sample(1:10, 5)
y <- sample(1:10, 5)
z <- sample(1:10, 5)

#2
A <- rbind(x, y, z)
A

#3
B <- cbind(x, y, z)
B

#4
m <- colMeans(A)
r <- rowMeans(A)
m
r

#5
colnames(B) <- c("V1", "V2", "V3")
B

```

## Subsetting matrix

1. Create a 5x5 matrix, `X`, sampled from 1:10 with replacement (`set.seed(50)`)
2. Create a 5x2 matrix, `Y` by selecting the 2nd and the 5th rows from `X`.
3. What are the numbers in row 2 and columns 3 and 5 in `X`
4. Select the 2nd column from `X` as a matrix, which is called a "column vector" in linear algebra

```{r m3, exercise=TRUE}
#1
set.seed(50)
X <- matrix(sample(1:10, 25, replace = TRUE), 5, 5)
X

#2
Y <- cbind(X[2,], X[5,])
Y

#3
X[2, c(3,5)]

#4
X[, 2, drop = FALSE]
```

1. What are the numbers in a row when the number in column 3 is 1?
2. Change 10's to 0's
3. Change 0's in the 5th column to 11

```{r m4-setup}
set.seed(50)
X <- matrix(sample(1:10, 25, replace = T), 5,5)
```

```{r m4, exercise=TRUE}
#1
X[X[,3]==1,]

#2
X[X==10] <- 0
X

#3
X[X[,5]==0, 5] <- 11
X
```

1. Find the index (row, column) where 2 is located in `X`.  Hint use `which()`
2. Multiply it with 0.75

```{r m5-setup}
set.seed(50)
X <- matrix(sample(1:10, 25, replace = T), 5,5)
```

```{r m5, exercise=TRUE}
#1
ind <- which(X == 2, arr.ind = TRUE)
ind

#2
X[ind] <- X[ind]*0.75
X
```

Create a 6 × 10 matrix of random integers chosen from 1, 2,. . . ,10 with replacement.  Call it `aMat` (with `set.seed(75)`)
  
1. Find the number of entries in each row which are greater than 4. You can use `apply()` function. 
2. Which row(s) contain exactly two occurrences of 7, if any? 

```{r m6, exercise=TRUE}
set.seed(75)
aMat <- matrix( sample(10, size=60, replace=T), nr=6)
aMat

#1
apply(aMat, 1, function(x) sum(x>4))

#2
which(apply(aMat, 1, function(x) sum(x==7)==2))
```

Write a script to find row and column index of each maximum and minimum values in `M`.

```{r}
set.seed(143)
M = matrix(sample(c(1:10), 16, replace = TRUE), nrow = 4, byrow = TRUE)
M
```

```{r m6b, exercise=TRUE}
which(M == max(M), arr.ind=TRUE)
which(M == min(M), arr.ind=TRUE)
```

## Dataframes

Dataframes are the most common way to store and interact with data. Data sets for statistical analysis are typically stored in dataframes in R. Unlike a matrix, a data frame can have different data types for each elements (columns).  We will use `whale` in your `RBootcamp`.  To see the description of the data call `?whale` in the console.

Use `str()`, `dim()`, `head()`, and `summary()` functions to examine the data

```{r m7, exercise=TRUE}
str(whale)
dim(whale)
head(whale)
summary(whale)
```

Extract all the elements of the first 8 rows and the last 4 columns of the `whale` dataframe 

```{r m8, exercise=TRUE}
nc <- ncol(whale)
rm <- nc - 4
whale[1:8, rm:nc]
```

Extract rows 9, 11, 25 and all columns except for the last column. 

```{r m9, exercise=TRUE}
whale[c(9, 11, 25), -ncol(whale)]
```

Now, we will extract data based on the following criteria, save the results:

1. At `depth` less than 1000m
2. `gradient` steeper than 150 degrees
3. All the observations less than the mean `depth`
4. `water.noise` level of "low"
5. `water.noise` level of "high" in the month of "May"
6. `month` of "October", `water.noise` level of "low" and `gradient` greater than the median value. Hint use `median()`
7. All observations between `latitude`s 60.0 and 61.0 and `longitude`s -6.0 and -4.0
8. All rows that do not have a `water.noise` level of "medium"

```{r m10, exercise=TRUE}
#1
whaled <- whale[whale$depth > 1000, ]
#2
whaleg <- whale[whale$gradient > 150, ]
#3
whaleall <- whale[whale$depth > mean(whale$depth), ]
#4
whalelow <- whale[whale$water.noise == "low", ]
#5
whalehmay <- whale[whale$water.noise == "high" & whale$month == "May", ]
#6
whalesub <- whale[whale$month == "October"
                  & whale$water.noise == "low"
                  & whale$gradient > median(whale$gradient), ]

whalelatlong <- whale[whale$latitude > 60
                        & whale$latitude < 61
                        & whale$longitude > -6
                        & whale$longitude < -4, ]

whalenomed <- whale[whale$water.noise != "medium", ]
```

There are of course many other approaches to handle these selection operations. One such approach is to use the `subset()` function (see `?subset`).  Here is an example for #5 given above:

```{r m11, exercise=TRUE}
subset(whale, month == "October"
       & water.noise == "low"
       & gradient > median(gradient))
```

Use `subset()` to extract data collected in "October" from `latitude`s greater than 61 degrees but only include the columns `month`, `latitude`, `longitude` and `number.whales`.

```{r m12, exercise=TRUE}
subset(whale, month == "October" & latitude > 61,
       select = c("month", "latitude", "longitude", "number.whales"))
```


