---
title: "Ch's 1&2 - Getting to know R and RStudio & Vectors"
author: "**Yigit Aydede - RBootcamp - 2022**"
output:
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(nycflights13)
library(Lahman)
tutorial_options(
  exercise.timelimit = 60,
  # A simple checker function that just returns the message in the check chunk
  exercise.checker = function(check_code, ...) {
    list(
      message = eval(parse(text = check_code)),
      correct = logical(0),
      type = "info",
      location = "append"
    )
  }
)
knitr::opts_chunk$set(error = TRUE)
```



## Set up

Create a new RStudio Project (select `File –> New Project` on the main menu). Create the Project in a new directory by selecting ‘New Directory’ and then select `New Project`. Give the Project a proper name in the `Directory name:`  Choose where you would like to create this Project directory. This will be your main RStudio Project file and directory which you will be using throughout this Rbootcamp.

## Script

Now create a new R script inside this Project from the main menu. Before you start putting codes, save this script. Call this script `exercise_1`. Remember, files names should be free of spaces. Check RStudio to see whether your file has been saved in the correct location.
  
At the top of your script you should include some information (metadata) to help your collaborators (and you) about the script by using `#` at the beginning of every line.

Practice the following codes in the script. 

1. Write `plot(1:10)` the execute it.
2. Display the help file for the function `mean`.
3. Create a variable called `my_first` and assign it the value 42.  Do you see in the ‘Environment’ tab in the top right window?
4. Create a variable called `my_second` and assign it the value 43.
5. Assign 45 to `my_first`.  Did value change?
6. Assign `my_first` to `my_second`.
7. Remove `my_second`.
8. Save R script.  Check it if it's saved or not.

## Calculator

Write the R code required to add two plus two:

```{r two-plus-two, exercise=TRUE}

```

Calculate $e^x \cos (x)$ at $x=3,3.1,3.2$

```{r expo, exercise=TRUE}

```

Find the natural log, log to the base 10, log to the base 2, square root and the natural antilog of 12.43. 

```{r log, exercise=TRUE}

```

## Vectors

Many operations in R make heavy use of vectors. Possibly the most common way to create a vector in R is using the `c()` **function**, which is short for “combine.” As the name suggests, it combines a list of elements separated by commas.

### Built-in functions

The base installation of R comes with many functions.  However, the power of R by comes from thousands of packages freely available. Moreover, once we get a bit more experience with using R you may want to define your own functions to perform tasks that are specific to your goals (more about this in Chapter 7).

The first function we will learn about is the `c()` function, which is short for concatenate and we use it to join together a series of values and store them in a data structure called a vector (more on vectors in Chapter 3).

Create a text vector called `Months` with names of the 12 months of the year.

```{r exp0, exercise=TRUE, exercise.eval=FALSE}

```

Now suppose we have 

```{r}
weight <- c(69, 62, 57, 59, 59, 64, 56, 66, 67, 66)
height <- c(112, 102, 83, 84, 99, 90, 77, 112, 133, 112, 110)
age <- c(12, 11, 13, 13, 13, 10, 11, 12, 14, 10, 11)
names <- c("a", "b", "c", "d", "e", "f", "g", "h", "i", "n")
```

```{r exp1-setup}
weight <- c(69, 62, 57, 59, 59, 64, 56, 66, 67, 66)
height <- c(112, 102, 83, 84, 99, 90, 77, 112, 133, 112)
age <- c(12, 11, 13, 13, 13, 10, 11, 12, 14, 10, 11)
names <- c("a", "b", "c", "d", "e", "f", "g", "h", "i", "n")
```

1. Calculate the body mass index (BMI) for each child. $\text{BMI} = weight/height^2$.
2. Combine `weight` and `height` into vector `x_com`,
3. Combine `eight`, `height`, and `names` into vector `x_all`,
4. Add `age` to `weight` into vector `x_add`

```{r exp1, exercise=TRUE, exercise.eval=FALSE}

```

Using `height`, calculate the mean, variance, standard deviation and number of elements in our vector. 

```{r exp2, exercise=TRUE, exercise.eval=FALSE}

```

### Patterns

Create a vector of the values of $e^x \cos (x)$ at $x=3,3.1,3.2, \ldots, 6$

```{r exp3, exercise=TRUE}

```

Create a vector $(1,2,3, \ldots, 19,20)$ and assign it to `z`

```{r exp4, exercise=TRUE}

```

Create a vector $(4,6,3,4,6,3, \ldots, 4,6,3)$ where there are 10 occurrences of 4. Assign it to `z`

```{r exp5, exercise=TRUE}

```

Calculate the following: $\sum_{i=10}^{100}\left(i^{3}+4 i^{2}\right)$ and assign it to `z`. 

```{r exp6, exercise=TRUE}

```

### Character

Concatenate it into a single sentence
```{r exp7, exercise=TRUE}
wr <- c("we have", "dogs", "cats", "and, donkey")
```

Use `paste()` to create the following character vectors of length 30 : ("label 1", "label 2", "label 30"). Note that there is a single space between label and the number following and you need to create fist the character vecrtor.

```{r exp8, exercise=TRUE}

```

### Sort, rank & order

The following lines created two vectors of random integers which are chosen with replacement from the integers $0, 1, … , 999$. Both vectors have length 250.

```{r}
set.seed(50)
xVec <- sample(0:999, 250, replace=T)
yVec <- sample(0:999, 250, replace=T)
```

```{r exp9-setup}
set.seed(50)
xVec <- sample(0:999, 250, replace=T)
yVec <- sample(0:999, 250, replace=T)
```

1. Sort the numbers in vector `xVec` by decreasing order
2. Find the order of values in vector `yVec`

```{r exp9, exercise=TRUE}

```

### Decriptives

Let's have 1000 values drawn from a normal distribution:

```{r}
set.seed(123)
h <- rnorm(1000)
```

```{r exp10-setup}
set.seed(123)
h <- rnorm(1000, mean = 5, sd = 3)
```

1. Find its summary statistics
2. What's the minimum and maximum values?
3. Make it standard normal distribution:  subtract its mean from each value and divide them with the standard error

```{r exp10, exercise=TRUE}

```

### Subsetting

If you want to extract particular elements from a vector, the extracted elements will give you a new vector. This process is called vector subsetting

1. Create a text vector called Months with names of the 12 months of the year.
2. Create a numeric vector "Summer", with Calendar month index positions for the summer months (inclusive, 4 total)
3. Use vector indexing to extract the text values of Months, indexed by Summer

```{r exp11, exercise=TRUE}

```

Let's have a simple vector with 10 values:

```{r}
set.seed(122)
h <- sample(1:10, replace = TRUE)
```

```{r exp12-setup}
set.seed(12)
h <- sample(1:10, replace = TRUE)
```

1. Extract 1, 5, and 7th elements in `h` and assign it to `a`
2. Find the values larger than 2 and assign it to a new vector `y`
3. Subtract the elements whose values are 5 and assign the remaining vector to `b`
4. Replace 2's with zero assign the new vector to `d`
5. Which element is maximum?
6. Which element is minimum?
7. Change 2's in `h` by multipying them with 0.7

```{r exp12, exercise=TRUE}

```

We have again the same two vectors:

```{r}
set.seed(50)
xVec <- sample(0:999, 250, replace=T)
yVec <- sample(0:999, 250, replace=T)
```

```{r exp13-setup}
set.seed(50)
xVec <- sample(0:999, 250, replace=T)
yVec <- sample(0:999, 250, replace=T)
```

Sort the numbers in the vector `xVec` in the order of increasing values in `yVec`

```{r exp13, exercise=TRUE}

```
  
### Missing values

You have a vector called `x`, without calling it:

```{r exp14-setup}
set.seed(150)
x <- sample(0:999, 1000, replace=TRUE)
NAs <- sample(1:900, 200, replace = FALSE)
x[NAs] <- NA
```

1. Do you have any NA's in `x?  Write a code! 
2. How many NA's do you have?  Write a code!
3. Which values are NA's?
4. Calculate the mean of `x`
5. Remove NA's from `x`


```{r exp14, exercise=TRUE}

```

